%\section{Lecture -- Week 6}
%\label{6}

\section{Group algebras}

Fix a field $K$ (for example, $K=\C$). 
For a finite group $G$, let $K[G]$ be the vector space (over $K$)
with basis $\{g:g\in G\}$. 
Thus $K[G]$ is a vector space of dimension $\dim K[G]=|G|$.
Moreover, $K[G]$ is a ring
with
\[
\left(\sum_{g\in G}\lambda_gg\right)\left(\sum_{h\in G}\mu_hh\right)
=\sum_{g,h\in G}\lambda_g\mu_h(gh).
\] 
Note that $K[G]$ is commutative if and only if $G$ is abelian. 

\begin{example}
    Let 
    \[
    \Sym_3=\{\id,(12),(13),(23),(123),(132)\}
    \]
    the symmetric group in three letters. Every element of $\C[\Sym_3]$
    is of the form 
    \[
    a\id+b(12)+c(13)+d(23)+e(123)+f(132)
    \]
    for some $a,b,c,d,e,f\in\C$. For example, 
    \[
    \alpha=5\id +3(123)\quad\text{ and }\quad 
    \beta=-4\id+(132)
    \]
    are elements of
    $\C[\Sym_3]$. We compute
    \begin{align*}
    \alpha+\beta &=1\id+3(123)+(132)
    \shortintertext{and}
    \alpha\beta &=(5\id +3(123))(-4\id+(132))\\
    &=-20\id +5(132)-12(123)+3(123)(132)\\
    &=-20\id +5(132)-12(123)+3\id\\
    &=-17\id+ 5(132)-12(123).
    \end{align*}
    Another example:
    \begin{align*}
    (\id+5(13))(2\id-(12))
    &=2\id-(12)+10(13)-5(13)(12)\\
    &=2\id-(12)+10(13)-5(123),
    \end{align*}
    as $(13)(12)=(123)$. 
\end{example}

Note that $K[G]$ is a ring and also a vector space (over $K$) and these structures
are somewhat compatible, as the formulas 
\[
(\lambda a+\mu b)c=\lambda (ac)+\mu (bc),\quad
a(\lambda b+\mu c)=\lambda (ab)+\mu (ac)
\]
hold for all $\lambda,\mu\in K$ and $a,b,c\in K[G]$. 

\begin{definition}
\index{Algebra}
Let $A$ be a ring and $K$ be a field. Then $A$ is a \emph{$K$-algebra} (or an \emph{algebra} 
over the field $K$) if $A$ is a vector space (over $K$)
such that $\lambda(ab)=(\lambda a)b=a(\lambda b)$ for all $\lambda\in K$ and $a,b\in A$. 
\end{definition}

The fields $\R$ and $\C$ are examples of $\R$-algebras. 
Other examples of $\R$-algebras are the 
polynomial rings $\R[X]$ and $\R[X,Y]$ and the matrix rings $M_n(\R)$.  

\begin{example}
	If $A$ is an algebra, then $M_n(A)$ is an algebra.	
\end{example}

\begin{example}
    Let $G$ be a finite group and $K$ be a field. Then $K[G]$ is a $K$-algebra. 
\end{example}

\begin{example}
	Let $G=\langle g:g^3=1\rangle=\{1,g,g^2\}\simeq C_3$ be the cyclic group of order three. 
	If $\alpha=a_11+_2g+a_3g^2$ and $\beta=b_11+b_2g+b_3g^2$, then
	\[
		\alpha\beta=(a_1b_1+a_2b_3+a_3b_2)1+(a_1b_2+a_2b_1+a_3b_3)g+a_1b_3+a_2b_2+a_3b_1)g^2.
	\]
	One can check that $\C[G]\simeq\C[X]/(X^3-1)$. 
\end{example}

In general, one proves that the group algebra of $C_n$, the cyclic group of order $n\geq2$, 
is isomorphic to $\C[X]/(X^n-1)$.

\begin{exercise}
\label{xca:RC3}
	Prove that $\R[C_3]\simeq\R\times\C$. 	
\end{exercise}

\begin{exercise}
	Let $G=\{1,g\}\simeq C_2$ be the cyclic group of order two. The product
	of $\C[G]$ is 
	\[
	(a1+bg)(c1+gd)=(ac+bd)1+(ad+bc)g.
	\]
	Prove that the map 
 \[
 \C[G]\to \C\times \C,\quad a1+bg\mapsto (a+b,a-b),
 \]
	is a linear isomorphism of rings. 
\end{exercise}

\begin{exercise}
	Let $K=\Z/2$ and $G=\{1,g\}\simeq C_2$ be the cyclic group of order two. 
	Prove that the map 
 \[
 K[G]\to\begin{pmatrix}
		K&K\\
		0&K
	\end{pmatrix},\quad 
 a1+bg\mapsto\begin{pmatrix}
		a+b&b\\
		0&a+b		
	\end{pmatrix},
    \]
    is a linear isomorphism of rings.
\end{exercise}

The group ring has the following property, which is left as an exercise. 
Let $A$ be an algebra and
$G$ be a finite group. If $f\colon G\to\mathcal{U}(A)$ is a group homomorphism, 
then there exists a unique algebra homomorphism $\varphi\colon K[G]\to A$ such that
$\varphi|_G=f$. 

\begin{example}
	Let $\D_3=\langle r,s:r^3=s^2=1,\,srs^{-1}=r^{-1}\rangle$ be the dihedral
	group of six elements. We claim that 
	\[
	\C[\D_3]\simeq\C\times\C\times M_2(\C).
	\]
	Let $\omega$ be a primitive cubic root of one. Let 
	\[
	R=\begin{pmatrix}
		\omega&0\\
		0&\omega^2	
	\end{pmatrix},
	\quad
	S=\begin{pmatrix}
		0&1\\
		1&0
	\end{pmatrix}.
 	\]
 	One easily checks that $SRS^{-1}=R^{-1}$ and $R^3=S^2=\begin{pmatrix}
		1&0\\
		0&1	
	\end{pmatrix}$. It follows that there exists a group homomorphism
	$G\to\C\times\C\times M_2(\C)$ such that
	$r\mapsto (1,1,R)$ and $s\mapsto (1,-1,S)$. This group homomorphism
	is a ring isomorphism.  
\end{example}


\section{Group representations}

\begin{definition}
	\index{Representation}
	A \emph{representation} (over the field $K$) of a group $G$ is a group homomorphism
	$\rho\colon G\to\GL(V)$, $g\mapsto\rho_g$, for some non-zero finite-dimensional 
        vector space $V$ (over $K$).
\end{definition}

\index{Representation!degree}
The \emph{degree} of the representation $\rho\colon G\to\GL(V)$ will be the dimension of $V$. 

Note that a representation $\rho\colon G\to\GL(V)$
satisfies the following properties:
\begin{enumerate}
    \item $\rho_{1_G}=\id$.
    \item $\rho_{gh}=\rho_g\rho_h$ for all $g,h\in G$.
    \item $\rho_g(v+w)=\rho_g(v)+\rho_g(w)$ for all $v,w\in V$.
    \item $\rho_g(\lambda v)=\lambda \rho_g(v)$ for all $v\in V$ and $\lambda\in K$. 
\end{enumerate}

Since by assumption
$V$ is finite-dimensional, say $\dim V=n$, after fixing a basis of $V$ we get a \emph{matrix representation} 
$\rho\colon G\to\GL(V)\simeq\GL_n(K)$ of $G$. Depending on the context, we will use either representations on vector spaces or matrix representations.  

\begin{example}
	We use group representations to show that the group 
	\[
    G=\langle x,y:x^2=y^2=1\rangle
    \]
    is infinite. Note that
	\[
	\rho\colon G\to\GL_2(\C),
	\quad
	x\mapsto\begin{pmatrix}
		1&0\\
		0&-1	
	\end{pmatrix},\quad
 	y\mapsto\begin{pmatrix}
		1&1\\
		0&-1	
	\end{pmatrix},
 	\]
 	is a group homomorphism, as 
 	$\rho_x^2=\rho_y^2=\begin{pmatrix}
		1&0\\
		0&1	
	\end{pmatrix}$. We claim that the elements of the form $(xy)^n$ are
	different for all $n$. It is enough to show that   
	$(xy)^n=(xy)^m$, then $n=m$. Note that
	\[
	\rho_{xy}=\rho_x\rho_y=\begin{pmatrix}
		1&0\\
		0&-1	
	\end{pmatrix}
	\begin{pmatrix}
		1&1\\
		0&-1	
	\end{pmatrix}
	=\begin{pmatrix}
		1&1\\
		0&1	
	\end{pmatrix}.
	\]
	Thus $\rho_{xy}^n=\begin{pmatrix}
		1&n\\
		0&1	
	\end{pmatrix}$ and 
	$\rho_{xy}^m=\begin{pmatrix}
		1&m\\
		0&1	
	\end{pmatrix}$. Then 
	\[
	(xy)^n=(xy)^m\implies\rho_{xy}^n=\rho_{xy}^m\implies n=m.
	\] 
\end{example}

The previous example demonstrates the power of group representations, even for infinite groups. 
However, in this course, we will focus on complex finite-dimensional representations of finite groups.

\begin{convention}
    When studying representations of groups, 
    we will focus on complex representations of finite groups, unless stated otherwise.
\end{convention}

\begin{example}
\index{Representation!trivial}
	If $G$ is a group, then $\rho\colon G\to\GL_1(\C)=\C^\times$, $g\mapsto 1$, 
	is a representation. This representation is known as the \emph{trivial (complex) representation} of $G$. 	
\end{example}

\begin{example}
	The sign of permutations yields a representation of $\Sym_n$. It is the group homomorphism
	$\Sym_n\to\C^\times$, $\sigma\mapsto\sgn(\sigma)$.  	
\end{example}

\begin{example}
	Let $G=\langle g:g^6=1\rangle$ be the cyclic group of order six. Then
	\[
	\rho\colon G\to\GL_2(\C),
	\quad 
	g\mapsto\begin{pmatrix}1&1\\-1&0\end{pmatrix}
	\] 
	is a group representation of degree two. 
\end{example}

%To prove the following result we need
%the \emph{minimal polynomial characterization of diagonazability",  
%see for example, \cite[Corollary 2.5.12]{zbMATH06049469}.
%
%\begin{proposition}
%    Let $G$ be a finite group and $\rho\colon G\to\GL_n(\C)$ 
%    be a representation. Then
%    each $\rho_g$ is diagonalizable.
%\end{proposition}
%
%\begin{proof}
%    Since $g$ is finite, $g^m=1$ for some $m\in\Z_{>0}$. This means that $\rho_g$ is a root of $X^m-1\in\C[X]$. Since
%    the roots of the polynomial $X^m-1$ are all different and $X^m-1$ factorizes linearly on $\C[X]$, it follows
%    that the minimal polynomial of $\rho_g$ also factorizes linearly in $\C[X]$. Hence $\rho_g$ is diagonalizable.
%\end{proof}
%
%In page \pageref{rho_diagonalizable} we will see an alternative  
%proof of the previous proposition. 
%
\begin{definition}
	\index{Invariant map}
    Let $\rho\colon G\to\GL(V)$ and $\psi\colon G\to\GL(W)$ be
    representations. A linear map $T\colon V\to W$ is said to be \emph{invariant} 
    if the diagram
    \[\begin{tikzcd}
	V & V \\
	W & W
	\arrow["{\rho_g}", from=1-1, to=1-2]
	\arrow["T"', from=1-1, to=2-1]
	\arrow["{\psi_g}"', from=2-1, to=2-2]
	\arrow["T", from=1-2, to=2-2]
\end{tikzcd}
\]
    commutes for all $g\in G$, this means that 
    $\psi_gT=T\rho_g$ for all $g\in G$.
\end{definition}

It is convenient to introduce an alternative
notation for group representations. 
Let 
\[ 
\rho\colon G\to\GL(V),\quad g\mapsto \rho_g,
\]
be a representation. For
$g\in G$ and $v\in V$ write $g\cdot v=\rho_g(v)$. Then the following properties hold: 
\begin{enumerate}
	\item $1\cdot v=v$ for all $v\in V$, 
	\item $g\cdot (h\cdot v)=(gh)\cdot v$ for all $g,h\in G$ and $v\in V$, 
	\item $g\cdot (v+w)=g\cdot v+g\cdot w$ for all $g\in G$ and $v,w\in V$, and 
	\item $g\cdot (\lambda v)=\lambda (g\cdot v)$ for all $g\in G$, $\lambda\in\C$ and $v\in V$.  	
\end{enumerate}
Using this notation for the representations $\rho\colon G\to\GL(V)$ 
and $\psi\colon G\to\GL(W)$, a linear
map $T\colon V\to W$ is invariant if and only if
\[
T(g\cdot v)=g\cdot T(v)
\]
for all $v\in V$ and $g\in G$. 
%Although we use the same notation for different representations, there is no risk of confusion. 

\begin{definition}
	\index{Representations!equivalence}
    The representations $\rho\colon G\to\GL(V)$ and $\psi\colon G\to\GL(W)$ are \emph{equivalent}
    if there exists a bijective linear 
    map $T\colon V\to W$ invariant. 
\end{definition}

If the representations $\rho$ and 
$\psi$ are equivalent, we write 
$\rho\simeq\psi$. 

\begin{example}
    Let $G=\Z/n$. The representations
    \begin{align*}
    &\rho\colon G\to\GL_2(\C),\quad
    m\mapsto
    \begin{pmatrix}
        \cos(2\pi m/n) & -\sin(2\pi m/n)\\
        \sin(2\pi m/n) & \cos(2\pi m/n)
    \end{pmatrix},
    \shortintertext{and}
    &\psi\colon G\to\GL_2(\C),\quad
    m\mapsto
    \begin{pmatrix}
        e^{2\pi im/n} & 0\\
        0 & e^{-2\pi im/n}
    \end{pmatrix},
    \end{align*}
    are equivalent, as $\rho_mT=T\psi_m$ for all $m\in G$ if $T=\begin{pmatrix}
        i&-i\\
        1&1
    \end{pmatrix}$.
\end{example}

The following example explains why we can alternatively use
representations or matrix representations. 

\begin{example}
\label{xca:change_of_basis}
Let $\rho\colon G\to\GL(V)$ be a representation, say, of degree 
$n=\dim V$. Choose a basis 
$B=\{v_1,\dots,v_n\}$ of $V$ and let 
\[
T\colon V\to\C^n,
\quad
\sum_{i=1}^n\lambda_iv_i\mapsto (\lambda_1,\dots,\lambda_n),
\]
be the isomorphism that takes coordinates 
for the basis $B$. For each $g\in G$, 
the composition of linear maps 
% https://q.uiver.app/#q=WzAsNCxbMCwwLCJcXENebiJdLFsyLDAsIlYiXSxbMSwwLCJWIl0sWzMsMCwiXFxDXm4iXSxbMCwyLCJUXnstMX0iXSxbMiwxLCJcXHJob19nIl0sWzEsMywiVCJdLFswLDMsIiIsMix7ImN1cnZlIjozfV1d
\[\begin{tikzcd}
	{\C^n} & V & V & {\C^n}
	\arrow["{T^{-1}}", from=1-1, to=1-2]
	\arrow["{\rho_g}", from=1-2, to=1-3]
	\arrow["T", from=1-3, to=1-4]
\end{tikzcd}
\]
produces a representation 
$\psi\colon G\to\GL(C^n)$
of $G$ equivalent to $\rho$.   
\end{example}



\begin{definition}
	\index{Invariant!subspace}
    Let $\rho\colon G\to\GL(V)$ be a representation. 
    A subspace $W$ of $V$ is said to be \emph{invariant} (with respect to $\rho$)
    if $\rho_g(W)\subseteq W$ for all $g\in G$.
\end{definition}

Let $\rho\colon G\to\GL(V)$ be a representation. Then $\{0\}$ and $V$ are always   
invariant subspaces of~$V$. Moreover, if $W\subseteq V$ is invariant, then
the map $\rho|_W\colon G\to\GL(W)$, $g\mapsto (\rho_g)|_W$, is a representation. The
map $\rho|_W$ is the \emph{restriction} of $\rho$ to $W$. 

\begin{exercise}
\index{Invariant map!kernel}
\index{Invariant map!image}
	Let $\rho\colon G\to\GL(V)$ and $\psi\colon G\to\GL(W)$ be representations of a given group $G$ 
	and 
 $T\colon V\to W$ be an invariant map. Prove that the \emph{kernel} 
	\[
	\ker T=\{v\in V:T(v)=0\}
	\]
	is an invariant subspace of $V$ and that the \emph{image} 
	\[
		T(V)=\{T(v):v\in V\}
	\]
	is an invariant subspace of $W$. 
\end{exercise}


\begin{definition}
\index{Subrepresentation}
    Let $\rho\colon G\to\GL(V)$
    be a representation. A \emph{subrepresentation} of $\rho$ is a restricted representation 
    of the form $\rho|_W\colon G\to\GL(W)$ for some invariant subspace $W$ of~$V$.
\end{definition}

%\begin{example}
%    Let $G=\langle g:g^3=1\rangle$ be the
%    cyclic group of order three
%    and
%    \[
%    \rho\colon G\to\GL_3(\R),
%    \quad
%    g\mapsto\begin{pmatrix}
%        0&1&0\\
%        0&0&1\\
%        1&0&0
%    \end{pmatrix}.
%    \]
%    The subspace
%    \[
%    W=\left\{
%    \begin{pmatrix}
%    x\\
%    y\\
%    z
%    \end{pmatrix}\in\R^3:x+y+z=0\right\}
%    \]
%    is a invariant subspace of $\R^3$.
%\end{example}

\begin{definition}
    \index{Representation!irreducible}
    A non-zero representation $\rho\colon G\to\GL(V)$ is \emph{irreducible} if
    $\{0\}$ and $V$ are the only invariant subspaces of $V$.
\end{definition}

Degree-one representations are irreducible. 

The irreducibility 
of a representation, of course, depends on the base field. 
In the following example, it is crucial using real representations.

\begin{example}
    Let $G=\langle g:g^3=1\rangle$ be the
    cyclic group of order three
    and
    \[
    \rho\colon G\to\GL_3(\R),
    \quad
    g\mapsto\begin{pmatrix}
        0&1&0\\
        0&0&1\\
        1&0&0
    \end{pmatrix}.
    \]
    Note that in this example we work with a real representation! 
    It is a routine calculation to prove that 
    \[
    W=\left\{
    \begin{pmatrix}
    x\\
    y\\
    z
    \end{pmatrix}\in\R^3:x+y+z=0\right\}
    \]
    is an invariant subspace of $\R^3$. We claim that $W$ 
    is irreducible. Let $S$ be a non-zero invariant subspace of 
    $W$ and let $s=\begin{pmatrix}x_0\\y_0\\z_0\end{pmatrix}\in S$ be a non-zero element. Then
    \[
    t=\begin{pmatrix}y_0\\z_0\\x_0\end{pmatrix}
    =\begin{pmatrix}
        0&1&0\\
        0&0&1\\
        1&0&0
    \end{pmatrix}
    \begin{pmatrix}x_0\\y_0\\z_0\end{pmatrix}\in S.
    \]
    We claim that $\{s,t\}$ is linearly independent. If not, there exists $\lambda\in\R$ such that
    $\lambda s=t$. Thus $\lambda x_0=y_0$, $\lambda y_0=z_0$ and $\lambda z_0=x_0$. This implies that
    $\lambda^3x_0=x_0$. Since $x_0\ne 0$ (because if $x_0=0$, then $y_0=z_0=0$, a contradiction), it follows that
    $\lambda=1$ and hence $x_0=y_0=z_0$, a contradiction because $x_0+y_0+z_0=0$.
    Therefore $\dim S=2$ and hence $S=W$.
\end{example}

What happens in the previous example if we work over complex numbers?

\begin{exercise}
    Let $\rho\colon G\to\GL(V)$ be a degree-two representation. Prove that
    $\rho$ is irreducible if and only if there is no common eigenvector for the $\rho_g$, $g\in G$.
\end{exercise}

The previous exercise can be used to show that the representation
$\Sym_3\to\GL_2(\C)$
of the symmetric group $\Sym_3$
given by
\[
(12)\mapsto\begin{pmatrix}
-1&-1\\0&1
\end{pmatrix},
\quad
(123)\mapsto\begin{pmatrix}
-1&-1\\
1&0
\end{pmatrix},
\]
is irreducible.

\begin{example}
\label{exa:direct_sum_reps}
Let $\rho\colon G\to\GL_n(\C)$ and $\psi\colon G\to\GL_m(\C)$ be representations. One defines
the \emph{direct sum} $\rho\oplus\psi$ of $\rho$ and $\psi$ as 
\[
\rho\oplus\psi\colon G\to\GL_{n+m}(\C),\quad
g\mapsto 
\begin{pmatrix}
\rho_g & 0\\
0 & \psi_g	
\end{pmatrix}.
\]
\end{example}

Let us describe the previous example without using matrix representations. If $V$ and 
$W$ are complex vector spaces, the (external) direct sum of $V$ and $W$ is defined
as the set $V\times W$ with the complex vector space structure given by
\[
(v,w)+(v_1,w_1)=(v+v_1,w+w_1),\quad
\lambda (v,w)=(\lambda v,\lambda w)
\]
for all $v,v_1\in V$, $w,w_1\in W$ and $\lambda\in\C$. 
If $\rho\colon G\to\GL(V)$ and $\psi\colon G\to\GL(W)$ are representations, 
the \emph{direct sum} $\rho\oplus\psi$ of $\rho$ and $\psi$ is then 
\[
\rho\oplus\psi\colon G\to\GL(V\oplus W),\quad
g\mapsto (\rho\oplus\psi)_g,
\]
where $(\rho\oplus\psi)_g\colon V\oplus W\to V\oplus W$, 
$(v,w)\mapsto (\rho_g(v),\psi_g(w))$. Note that both 
$V\simeq V\oplus\{0\}$ and $W\simeq \{0\}\oplus W$ are
invariant subspaces of $V\oplus W$. 

\begin{definition}
    \index{Representation!completely reducible}
    A representation $\rho\colon G\to\GL(V)$ is said to be 
    \emph{completely reducible}
    if $\rho$ can be decomposed as
    $\rho=\rho_1\oplus\cdots\oplus \rho_n$ for some irreducible
    representations $\rho_1,\dots,\rho_n$ of $G$. 
\end{definition}

Note that if $\rho\colon G\to\GL(V)$ is completely reducible and 
$\rho=\rho_1\oplus\cdots\oplus \rho_n$ for some irreducible representations 
$\rho_i\colon G\to\GL(V_i)$, $i\in\{1,\dots,n\}$, then 
each $V_i$ is an invariant subspace of $V$ and $V=V_1\oplus \cdots \oplus V_n$. 
Moreover, on some basis of $V$, the matrix  
$\rho_g$ can be written as 
\[
\rho_g=\begin{pmatrix}
(\rho_1)_g &  \\
& (\rho_2)_g  \\
&&\ddots\\
&&&(\rho_n)_g	
\end{pmatrix}.
\]

\begin{definition}
\index{Representation!decomposable}
\index{Representation!indecomposable}
A representation
$\rho\colon G\to\GL(V)$ is \emph{decomposable} if $V$ can be decomposed as $V=S\oplus T$
where $S$ and $T$ are non-zero invariant subspaces of $V$. 
\end{definition}

A representation is 
\emph{indecomposable} if it is not decomposable. 

\begin{exercise}
\label{xca:equivalence}
	Let $\rho\colon G\to\GL(V)$ and $\psi\colon G\to\GL(W)$ be equivalent representations.
	Prove the following facts:
	\begin{enumerate}
		\item If $\rho$ is irreducible, then $\psi$ is irreducible.
		\item If $\rho$ is decomposable, then $\psi$ is decomposable.
		\item If $\rho$ is completely reducible, then $\psi$ is completely reducible. 
	\end{enumerate}	
\end{exercise}

\section{Tensor product of representations}

Let $V$ be a vector space with basis $\{v_1,\dots,v_k\}$ and 
$W$ be a vector space with basis $\{w_1,\dots,w_l\}$. A 
\emph{tensor product} of $V$ and $W$ is a vector space $X$ with 
together with a bilinear map 
\[
V\times W\to X,
\quad
(v,w)\mapsto v\otimes w,
\]
such that $\{v_i\otimes w_j:1\leq i\leq k,\,1\leq j\leq l\}$ is a  
basis of $X$. The tensor product of $V$ and $W$ is unique up to isomorphism 
and is denoted by $V\otimes W$. Note that
\[
\dim(V\otimes W)=(\dim V)(\dim W).
\]
Moreover, every element of $V\otimes W$ is a finite sum 
of the form
\[
\sum_{i,j}\lambda_{ij}v_i\otimes v_j
\]
for some scalars $\lambda_{ij}\in\C$. It is important to remark that elements 
of $V\otimes W$ are not always of the form $v\otimes w$ for $v\in V$ and $w\in W$. 

The bilinearity of tensor products is crucial. 
For example,
\begin{align*}
    (v_1+v_3)\otimes (3w_1+w_2) 
    &=v_1\otimes (3w_1+w_2)+v_3\otimes (3w_1+w_2)\\
    &=v_1\otimes (3w_1)+v_1\otimes w_2+v_3\otimes (3w_1)+v_3\otimes w_2\\
    &=3v_1\otimes w_1+v_1\otimes w_2+3v_3\otimes w_1+v_3\otimes w_2.
\end{align*}

\begin{definition}
	Let $\rho\colon G\to\GL(V)$ and $\psi\colon G\to\GL(W)$ be representations, and 
    let $\{v_1,\dots,v_n\}$ be a basis of $V$ and $\{w_1,\dots,w_m\}$ a basis of $W$. 
    The \emph{tensor product} of $\rho$ and $\psi$ is the representation of $G$ given by 
	\begin{gather*}
	\rho\otimes\psi\colon G\to\GL(V\otimes W),
	\quad 
	g\mapsto (\rho\otimes\psi)_g,
	\shortintertext{where}
	(\rho\otimes\psi)_g\left (\sum_{i,j}\lambda_{ij} v_i\otimes w_j\right)=\sum_{i,j}\lambda_{ij} \rho_g(v_i)\otimes \psi_g(w_j),\quad 
    i\in\{1,\dots,n\},\, j\in\{1,\dots,m\}. 
	\end{gather*} 	
\end{definition}

\begin{exercise}
\label{xca:tensor}
Prove that the tensor product of representations 
is a representation.     
\end{exercise}



\section{Unitary representations}

Since we are considering finite-dimensional vector spaces, our vector spaces admit 
an \emph{inner product}, that is a map $V\times V\to\C$, $(v,w)\mapsto\langle v,w\rangle$, 
that satisfies the following properties:
\begin{enumerate}
    \item $\langle y,x\rangle=\overline{\langle x,y\rangle}$ for all $x,y\in V$, where $\overline{a+bi}=a-bi$ denotes the complex conjugate.
    \item $\langle ax+by,z\rangle=a\langle x,z\rangle+b\langle y,z\rangle$ for all $x,y,z\in V$ and $a,b\in\C$.
    \item $\langle x,x\rangle>0$ for all $x\in V$ and $\langle x,x\rangle=0$ if $x=0$. 
\end{enumerate}

For example, if $V=\C^n$, $v=\colvec{3}{v_1}{\vdots}{v_n}\in V$ and $w=\colvec{3}{w_1}{\vdots}{w_n}\in V$, 
then the \emph{usual inner product} between $v$ and $w$ 
is given by
\begin{equation}
    \label{eq:inner_product}
V\times V\to\C,\quad 
(v,w)\mapsto v_1\overline{w_1}+\cdots+v_n\overline{w_n}=v^T\overline{w},
\end{equation}
where 
\[
v^T=(v_1,\dots,v_n),\quad 
\overline{w}=\colvec{3}{\overline{w_1}}{\vdots}{\overline{w_n}}.
\]

\begin{definition}
    \index{Representation!unitary}
    Let $V$ be a finite-dimensional complex vector space with inner product $(v,w)\mapsto\langle v,w\rangle$. 
    A representation $\rho\colon G\to\GL(V)$ is \emph{unitary} if
    $\langle \rho_gv,\rho_gw\rangle=\langle v,w\rangle$ for all $g\in G$ and $v,w\in V$.
\end{definition}

\begin{exercise}
\index{Left regular representation}
Let $G$ be a finite group and $V=\C[G]$ be the complex vector space with basis
$\{g:g\in G\}$. The \emph{left regular representation}
of $G$ is the representation
\[
L\colon G\to\GL(V),
\quad
g\mapsto L_g,
\]
where $L_g(h)=gh$. With respect to the inner product
\[
\left\langle\sum_{g\in G}\lambda_gg,\sum_{g\in G}\mu_gg\right\rangle=\sum_{g\in G}\lambda_g\overline{\mu_g},
\]
the representation $L$ is unitary. Prove it! 
\end{exercise}



If $V$ is a vector space and $W$ is a subspace of $V$, 
\[
W^\perp = \{v\in V:\langle v,w\rangle=0\text{ for all $w\in W$}\}, 
\]
is called the 
\emph{orthogonal complement} of $W$. The following result is extremely important: 

\begin{proposition}
\label{pro:irr_or_dec}
Let $\rho\colon G\to\GL(V)$ be a unitary representation. Then $\rho$ is either
irreducible or decomposable.
\end{proposition}

\begin{proof}
	If $\rho$ is not irreducible, there exists an invariant subspace $W$ of $V$ such that
	$\{0\}\subsetneq W\subsetneq V$. Can we find an invariant complement? Yes, and this is because 
        our representation is unitary.  
	We know that $V=W\oplus W^{\perp}$ as complex vector spaces, where 
	$W^\perp\ne\{0\}$. So $\rho$ will be decomposable if we can prove that
	$W^\perp$ is an invariant subspace of $V$. 
	Let $v\in W^\perp$, $w\in W$ and $g\in G$. Then  
	\[
	\langle \rho_g(v),w\rangle=\langle v,\rho_g^{-1}(w)\rangle=0   
	\]
	since $\rho_g^{-1}(w)\in W$, as $W$ is an invariant subspace of $V$. This means that 
	$\rho_g(v)\in W^\perp$ and hence $V=W\oplus W^{\perp}$ is decomposable.  
\end{proof}

\section{Weyl's trick} 

The following theorem is known as the \emph{Weyl trick}. 

\begin{theorem}[Weyl]
\index{Weyl's!trick}
\index{Weyl's!theorem}
    Every representation $\rho\colon G\to\GL_n(\C)$ 
    of a finite group is unitary.
\end{theorem}

\begin{proof}
    Let $V=\C^n$ and 
    $V\times V\to\C$, $(v,w)\mapsto v\cdot w$, be an\footnote{It does not matter which inner product we have chosen, we can safely use, for example, the usual inner product of \eqref{eq:inner_product}.} inner
    product on $V$. 
   
    A straightforward calculation shows that
    \[
    \langle v,w\rangle=\sum_{g\in G}(\rho_gv)\cdot (\rho_gw)
    \]
    is an inner product of $V$. For example, 
    \begin{align*}
    \langle v,w+w_1\rangle&=\sum_{g\in G}\rho_gv\cdot\rho_g(w+w_1)\\
    &=\sum_{g\in G}\rho_gv\cdot \left(\rho_gw+\rho_gw_1\right)\\
    &=\sum_{g\in G}\rho_gv\cdot \rho_gw+\rho_gv\cdot\rho_gw_1\\
    &=\sum_{g\in G}\rho_gv\cdot \rho_gw+\sum_{g\in G}\rho_gv\cdot \rho_gw_1\\
    &=\langle v,w\rangle+\langle v,w_1\rangle 
    \end{align*}
    holds for all $v,w,w_1\in V$. 
    
    Since
    \begin{align*}
    \langle\rho_gv,\rho_gw\rangle&=\sum_{h\in G}(\rho_h\rho_gv)\cdot(\rho_h\rho_gw)\\
    &=\sum_{h\in G}(\rho_{hg}v)\cdot(\rho_{hg}w)=\sum_{x\in G}(\rho_xv)\cdot(\rho_xw)=\langle v,w\rangle,
    \end{align*}
    the representation $\rho$ is unitary.
\end{proof}

\begin{example}
    Let $A=\begin{pmatrix}
        -1 & -1\\
        -1 & 0
    \end{pmatrix}$. Since $A^3=I$, we can use this matrix to 
    represent the (multiplicative) cyclic group 
    \[
    C_3=\langle g:g^3=1\rangle
    \]
    of order three. Note that the map $1\mapsto g$ yields an isomorphism 
    $C_3\simeq \Z/3$. 
        
    The map $C_3\to\GL_2(\C)$, $g\mapsto A$, is a group homomorphism. In particular, we have a representation $\rho\colon C_3\to\GL(V)$, where $V=\C^2$ with the usual inner product given by \eqref{eq:inner_product}, that is 
    \[
    v\cdot w=v^T\overline{w}
    \]
    for $v,w\in V$. 
    For example, the inner product between the vectors $\colvec{2}{1}{i}$ and $\colvec{2}{1}{2}$
    is 
    \[
    (1,i)\colvec{2}{1}{2}=1+2i.
    \]
    Since $A\in M_2(\R)$, we compute 
    \[
    \langle v,w\rangle=v^T\overline{w}+(Av)^T\overline{(Aw)}+(A^2v)^T\overline{(A^2w)}=v^T\left (I+A^TA+(A^2)^TA^2\right)\overline{w}
    \]
    and this is an invariant inner product on $V$. Note that 
    \[
        B=I+A^TA+(A^2)^TA^2=\begin{pmatrix}
            4&2\\
            2&4
        \end{pmatrix}.
    \]
\end{example}

\label{rho_diagonalizable}
Weyl's trick has several interesting consequences:  

\begin{corollary}
\label{cor:consequences}
	Let $\rho\colon G\to\GL(V)$ be a representation of a finite group $G$. 
 The following properties
	hold:
	\begin{enumerate}
		\item $\rho$ is equivalent to a unitary representation.
		\item $\rho$ is either irreducible or decomposable.
		\item Each $\rho_g$ is diagonalizable. 
	\end{enumerate}
\end{corollary}

\begin{proof}
	The first claim follows from Example \ref{xca:change_of_basis} and 
	Weyl's trick. The second claim, from 1) and
	Proposition \ref{pro:irr_or_dec}. Finally, 3) follows 
	from Weyl's trick immediately, as unitary matrices are diagonalizable.  
\end{proof}

\begin{exercise}
\label{xca:not_decomposable}
    If $G$ is an infinite group, it is no longer true that every non-zero representation
    is either irreducible or decomposable. Find an example.
\end{exercise}

For the previous exercise. consider the representation $\Z\to\GL_2(\C)$, 
$1\mapsto\begin{pmatrix}1&1\\0&1\end{pmatrix}$. 
